MCTS Based on Simple Regret

### 2012-01-21 paper/srcr/paper.tex
moved to a new place

We introduce here two sampling schemes with superpolynomially
decreasing upper bounds on the simple regret. The bounds
suggest that these schemes achieve a lower simple regret
than uniform sampling; indeed, this is confirmed
by experiments.

@@

Bound the probability $P_j$ that a non-optimal arm $j$ is selected. Split the interval
 $[\mu_j, \mu_*]$ at $\mu_j+\delta_j$. By Chernoff-Hoeffding bound:
\begin{eqnarray}
P_j&\le&\Pr[\overline X_j>\mu_j+\delta_j]+\Pr[\overline X_*<\mu_*-(\Delta_j-\delta_j)]\nonumber\\
   &\le&\exp\left(-2\delta_j^2n_j\right)+\exp\left(-2(\Delta_j-\delta_j)^2n_*\right)
\end{eqnarray}
Observe that, in probability, $\overline X_j \rightarrow \mu_j$ as $n\rightarrow\infty$, 
therefore $n_*\rightarrow n\varepsilon$, $n_j\rightarrow\frac
{n(1-\varepsilon)} {K-1}$ as $n\rightarrow \infty$. Conclude that for
any $0<\pi<1$, $\gamma>1$ there exists $N$ such that for any $n>N$ and
all non-optimal arms $j$:
\begin{equation}
P_j \le \gamma \left(\exp\left(-\frac {2\delta_j^2n(1-\varepsilon)}{K-1}\right)
+\exp\left(-2(\Delta_j-\delta_j)^2n\varepsilon\right)\right)
\label{eqn:epsgreedy-probbound}
\end{equation}
Require
\begin{eqnarray}
\exp\left(-\frac {2\delta_j^2n(1-\varepsilon)}{K-1}\right)
&=&\exp\left(-2(\Delta_j-\delta_j)^2n\varepsilon\right)\nonumber\\
\frac {\delta_j} {\Delta_j-\delta_j}&=&\sqrt{\frac {(K-1)\varepsilon} {1-\varepsilon}}
\label{eqn:epsgreedy-constant-varepsilon}
\end{eqnarray}
Substitute (\ref{eqn:epsgreedy-probbound}) together with
(\ref{eqn:epsgreedy-constant-varepsilon}) into
(\ref{eqn:simple-regret}) and obtain
\begin{equation}
\IE r_{\varepsilon\mbox{-}greedy}\le 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {2\Delta_j^2n\varepsilon}
  {\left(1+\sqrt{\frac {(K-1)\varepsilon}
        {1-\varepsilon}}\right)^2}\right)
\end{equation}
\end{proof}

In particular, as the number of arms $K$ grows, the bound for $\frac 1
2$-greedy sampling ($\varepsilon=\frac 1 2$) becomes considerably tighter than for uniform
random sampling ($\varepsilon=\frac 1 K$):
\begin{crl}
For uniform random sampling, 
\begin{equation}
\IE r_{uniform}\le 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {\Delta_j^2n} {K}\right)
\end{equation}
For $\frac 1 2$-greedy sampling,
\begin{eqnarray}
\IE r_{\frac 1 2\mbox{-}greedy}&\le& 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {2\Delta_j^2n}
  {\left(1+\sqrt{(K-1)}\right)^2}\right)\nonumber\\
  &\approx& 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac
    {2\Delta_j^2n} {K}\right)\mbox{ for large K}\nonumber\\
\end{eqnarray}
\end{crl}

$\varepsilon$-greedy is based solely on sampling the empirically best
arm with a higher probability then the rest of the arms, and ignores
information about empirical means of other arms. On the other hand,
UCB distributes samples in accordance with empirical means, but, in order to
minimize cumulative regret, chooses the empirically best arm too often.
Intuitively, a better scheme for simple regret minimization would
distribute samples in a way similar to UCB, but would sample the best arm
less. This can be achieved by replacing $\log(\cdot)$ in
Equation~\ref{eqn:ucb} with a faster growing sublinear function, for
example, $\sqrt\cdot$.
\begin{dfn} Scheme $\mathbf{UCB_{\sqrt{\cdot}}(\alpha)}$ repeatedly pulls arm $i$ that
maximizes $b_i$:
\begin{equation}
b_i=\overline X_i+\sqrt {\frac {\alpha \sqrt n} {n_i}}
\end{equation}
where, as before, $\overline X_i$ is the average reward obtained from arm $i$,
$n_i$ is the number of times arm $i$ was pulled, and $n$ is the total
number of pulls so far. \end{dfn}
This scheme also exhibits a superpolynomially decreasing simple regret:
\begin{thm}  For any $0<\pi<1$  and $\gamma>1$ there exists $N$ such that for
  any number of samples $n>N$ the simple regret of the  UCB$_{\sqrt{\cdot}}$($\alpha$)
sampling scheme is  bounded from above as
\begin{equation}
\IE r_{ucb\sqrt{\cdot}} \le \gamma\sum_{i=1}^K\Delta_i\exp\left(-2\alpha\sqrt{n}\right)
\end{equation}
with probability at least $1-\pi$.
\end{thm}

\begin{proof}[Proof outline:] TODO
\end{proof}

>>>>

We introduce here two sampling schemes with super-polynomially
decreasing upper bounds on the simple regret. The bounds
suggest that these schemes achieve a lower simple regret
than uniform sampling; indeed, this is confirmed
by experiments.

@@

Bound the probability $P_i$ that a non-optimal arm $i$ is selected. Split the interval
 $[\mu_i, \mu_*]$ at $\mu_i+\delta_i$. By Chernoff-Hoeffding bound:
\begin{eqnarray}
P_i&\le&\Pr[\overline X_i>\mu_i+\delta_i]+\Pr[\overline X_*<\mu_*-(\Delta_i-\delta_i)]\nonumber\\
   &\le&\exp\left(-2\delta_i^2n_i\right)+\exp\left(-2(\Delta_i-\delta_i)^2n_*\right)
\end{eqnarray}
Observe that, in probability, $\overline X_i \rightarrow \mu_i$ as $n\rightarrow\infty$, 
therefore $n_*\rightarrow n\varepsilon$, $n_i\rightarrow\frac
{n(1-\varepsilon)} {K-1}$ as $n\rightarrow \infty$. Conclude that for
any $0<\pi<1$, $\gamma>1$ there exists $N$ such that for any $n>N$ and
all non-optimal arms $i$:
\begin{equation}
P_i \le \gamma \left(\exp\left(-\frac {2\delta_i^2n(1-\varepsilon)}{K-1}\right)
+\exp\left(-2(\Delta_i-\delta_i)^2n\varepsilon\right)\right)
\label{eqn:epsgreedy-probbound}
\end{equation}
Require
\begin{eqnarray}
\exp\left(-\frac {2\delta_i^2n(1-\varepsilon)}{K-1}\right)
&=&\exp\left(-2(\Delta_i-\delta_i)^2n\varepsilon\right)\nonumber\\
\frac {\delta_i} {\Delta_i-\delta_i}&=&\sqrt{\frac {(K-1)\varepsilon} {1-\varepsilon}}
\label{eqn:epsgreedy-constant-varepsilon}
\end{eqnarray}
Substitute (\ref{eqn:epsgreedy-probbound}) together with
(\ref{eqn:epsgreedy-constant-varepsilon}) into
(\ref{eqn:simple-regret}) and obtain
\begin{equation}
\IE r_{\varepsilon\mbox{-}greedy}\le 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {2\Delta_i^2n\varepsilon}
  {\left(1+\sqrt{\frac {(K-1)\varepsilon}
        {1-\varepsilon}}\right)^2}\right)
\end{equation}
\end{proof}

In particular, as the number of arms $K$ grows, the bound for $\frac 1
2$-greedy sampling ($\varepsilon=\frac 1 2$) becomes considerably tighter than for uniform
random sampling ($\varepsilon=\frac 1 K$):
\begin{crl}
For uniform random sampling, 
\begin{equation}
\IE r_{uniform}\le 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {\Delta_i^2n} {K}\right)
\end{equation}
For $\frac 1 2$-greedy sampling,
\begin{eqnarray}
\IE r_{\frac 1 2\mbox{-}greedy}&\le& 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {2\Delta_i^2n}
  {\left(1+\sqrt{(K-1)}\right)^2}\right)\nonumber\\
  &\approx& 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac
    {2\Delta_i^2n} {K}\right)\mbox{ for large K}\nonumber\\
\end{eqnarray}
\end{crl}

$\varepsilon$-greedy is based solely on sampling the empirically best
arm with a higher probability then the rest of the arms, and ignores
information about empirical means of other arms. On the other hand,
UCB distributes samples in accordance with empirical means, but, in order to
minimize cumulative regret, chooses the empirically best arm too often.
Intuitively, a better scheme for simple regret minimization would
distribute samples in a way similar to UCB, but would sample the best arm
less. This can be achieved by replacing $\log(\cdot)$ in
Equation~\ref{eqn:ucb} with a faster growing sublinear function, for
example, $\sqrt\cdot$.
\begin{dfn} Scheme $\mathbf{UCB_{\sqrt{\cdot}}(\alpha)}$ repeatedly pulls arm $i$ that
maximizes $b_i$:
\begin{equation}
b_i=\overline X_i+\sqrt {\frac {\alpha \sqrt n} {n_i}}
\end{equation}
where, as before, $\overline X_i$ is the average reward obtained from arm $i$,
$n_i$ is the number of times arm $i$ was pulled, and $n$ is the total
number of pulls so far. \end{dfn}
This scheme also exhibits a super-polynomially decreasing simple regret:
\begin{thm}  For any $0<\pi<1$  and $\gamma>1$ there exists $N$ such that for
  any number of samples $n>N$ the simple regret of the  UCB$_{\sqrt{\cdot}}$($\alpha$)
sampling scheme is  bounded from above as
\begin{equation}
\IE r_{ucb\sqrt{\cdot}} \le
2\gamma\sum_{i=1}^K\Delta_i\exp\left(-\frac {\alpha\sqrt{n}} 2\right)
\end{equation}
with probability at least $1-\pi$.
\end{thm}

\begin{proof}[Proof outline:] Bound the probability $P_i$ that a
  non-optimal arm $i$ is chosen. Split the interval $[\mu_i, \mu_*]$
  at $\mu_i+\frac {\Delta_i} 2$. By Chernoff-Hoeffding bound:
\begin{eqnarray}
P_i&\le&Pr\left[\overline X_i>\mu_i+\frac {\Delta_i} 2\right]+\Pr\left[\overline X_*<\mu_*-\frac {\Delta_i}
  2\right]\nonumber \\
   &\le&\exp\left(-\frac {\Delta_i^2n_i} 2\right)+\exp\left(-\frac {\Delta_i^2n_*} 2\right)
\end{eqnarray}
Observe that, in probability, $n_i\to \frac {\alpha \sqrt n}
{\Delta_i^2}$, $n_i\le n_*$ as $n\to\infty$. Conclude that for
any $0<\pi<1$, $\gamma>1$ there exists $N$ such that for any $n>N$ and
all non-optimal arms $i$:
\begin{equation}
P_i \le 2\gamma \exp\left(-\frac {\alpha \sqrt n} 2\right)
\label{eqn:usbsqrt-probbound}
\end{equation}
Substitute (\ref{eqn:usbsqrt-probbound}) into
(\ref{eqn:simple-regret}) and obtain
\begin{equation}
\IE r_{ucb\sqrt{\cdot}} \le 2\gamma\sum_{i=1}^K\Delta_i\exp\left(-\frac {\alpha\sqrt{n}} 2\right)
\end{equation}
\end{proof}
